%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

This dissertation would not have been possible without support from innumerable sources. 
First and foremost, I must thank Molly Babel, whose enthusiasm and passion for linguistic research is infectious.
Equally impressive is her focus and organization, which really helped get this dissertation done in (somewhat) timely fashion.
I could not imagine a better supervisor and collaborator.
Additionally, feedback and discussions from my committee members, Eric Vatikiotis-Bateson and Carla Hudson Kam, have been invaluable in shaping my thinking on all of my projects over the past five years.

Several seminars provided excellent perspectives on my dissertation research.
The first, taught by Martina Wiltschko and Michael Rochemont, helped shape my initial prospectus.
The second, taught by Molly Babel and Kathleen Currie Hall, provided a forum to share my intiial findings and thought process and helped refine them.
I am also grateful to my fellow students in those seminars for all the lively discussions.
My fellow graduate students have shaped my thinking and my graduate experience, namely: Kevin McMullin, Mark Scott, Anita Szakay, Jen Abel, Alexis Black, Heather Bliss, Blake Allen, Michael Fry, Scott Mackie, and Murray Schellenberg.  You're all awesome.

I would like to thank the members of the Speech in Context lab, who are uniformly wonderful people.  
In particular, Michelle Chan and Jobie Hui helped me with stimuli selection, and Jobie Hui and Jamie Russell helped me with running participants.
I am indebted to all of you.

My wonderful parents have supported me from the beginning of studying linguistics, even if it wasn't the most viable major for a career in undergrad.
I certainly would not have found my passion in linguistics if it weren't for the exposure to other languages and cultures that I had growing up.
My brother, aunts, uncles and cousins from around the world have been amazingly supportive as well.
Thank you especially to Laura Tammpere who has been a constant source of support and fun.
I love you all!